### 2016-03-17 text/MS.Rmd
added new paragraph about population-level energy use

1. Metabolic theory posits that size and temperature are fundamental drivers of ecological patterns and processes, via their respective associations with metabolic rates. Metabolic rates are well known to increase allometrically with body mass and approximately exponentially with temperature, but most applications of metabolic theory treat these effects as independent.  
2. Almost all studies of intraspecific scaling of unitary organisms must rely on ontogenetic differences in size to examine its relationship with metabolic rate, thereby potentially confounding age, nutrition or growth rate with size –- all of which can affect metabolic rate. In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.  
3. While increased temperature increased metabolic rate, we found no consistent effect of temperature or growth form on metabolic scaling; rather the effects were species-specific. We show that mass-scaling exponents and temperature dependencies interact in ways that are unanticipated by extant theory.  
4. The allometric scaling of metabolic rate that we observe in the present study contrasts with the isometric scaling of water clearance and feeding rates observed in some colonial marine invertebrates. Our findings suggest that the often observed competitive superiority of colonial organisms over unitary ones may arise because energy availability for growth and or reproduction (the difference between energy intake and expenditure) increases more strongly with size in colonial organisms.

@@

This differential between ingestion rates and metabolic costs may explain the observed exponential growth and reproduction observed in some colonial marine invertebrates (e.g. `r myCite(c('marshall2013evolution', 'padilla1996limn', 'winston2010icb'))`) -- because the ratio of intake to expenditure increases with size, larger colonies can allocate proportionately more energy to growth and/or reproduction. In other words, for these colonial marine invertebrates, larger colonies are more efficient at allocating food intake to somatic growth and/or reproduction than smaller colonies. Isometric energy gains and allometric energy expenditure in colonial marine invertebrates may explain why they are often competitively dominant, outgrowing and excluding unitary organisms (which generally present allometric scaling of both feeding and metabolic rates, see e.g. `r myCite(c('kooijman2009book', 'kearney2012amnat'))`) in a range of systems `r myCite(c('buss1980pnas', 'buss1990tree'))`.

@@

Here we have demonstrated that the effects of temperature interact in various ways with the mass-scaling of metabolic rates in different species of colonial marine invertebrates encompassing two distinct phyla. By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by mimicking predation, which constitutes a fundamental ecological process in the wild. Mass-scaling exponents vary for different species, and are not consistent across growth forms nor phylogeny. Finally, our study suggests that the often observed competitive superiority of colonial organisms over unitary ones may arise because energy availability for growth and or reproduction increases more strongly with size in colonial organisms.

@@

We thank M. R. Kearney for **what reason?**, field and laboratory assistants **names?**, the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C. R. White), and **whatever grants Dustin wants to add**.

>>>>

1. Metabolic theory posits that size and temperature are fundamental drivers of ecological patterns and processes, via their respective associations with metabolic rates. Metabolic rates tend to increase allometrically with body mass and approximately exponentially with temperature, but most applications of metabolic theory treat these effects as independent.  
2. Most studies of intraspecific scaling of unitary organisms must rely on ontogenetic differences in size to examine its relationship with metabolic rate, thereby potentially confounding age, nutrition or growth rate with size, all of which can affect metabolic rate. In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.  
3. While increased temperature increased metabolic rate, we found no consistent effect of temperature or growth form on metabolic scaling; rather the effects were species-specific. We show that mass-scaling exponents and temperature dependencies interact in ways that are unanticipated by extant theory.  
4. The allometric scaling of metabolic rate that we observe in the present study contrasts with the isometric scaling of water clearance and feeding rates observed in some colonial marine invertebrates. Our findings suggest that the often observed competitive superiority of colonial organisms over unitary ones may arise because energy availability for growth and or reproduction (the difference between energy intake and expenditure) increases more strongly with size in colonial organisms.

@@

This differential between ingestion rates and metabolic costs may explain the observed exponential growth and reproduction observed in some colonial marine invertebrates (e.g. `r myCite(c('marshall2013evolution', 'padilla1996limn', 'winston2010icb'))`) -- because the ratio of intake to expenditure increases with size, larger colonies can allocate proportionately more energy to growth and/or reproduction. In other words, for these colonial marine invertebrates, larger colonies are more efficient at garnering food and allocating it to somatic growth and/or reproduction than smaller colonies. Isometric energy gains and allometric energy expenditure in colonial marine invertebrates may explain why they are often competitively dominant, outgrowing and excluding unitary organisms (which generally present allometric scaling of both feeding and metabolic rates, see e.g. `r myCite(c('kooijman2009book', 'kearney2012amnat'))`) in a range of systems `r myCite(c('buss1980pnas', 'buss1990tree'))`.

@@

At the population level, disturbance and predation can also create different size distributions. Our finding of allometric scaling of metabolic rate implies non-intuitive outcomes of such shifts in mean size of colonies in a population. For example, at 25 ˚C a population of 10 colonies of *Bugula neritina* that each weigh 10 mg would have a cumulative metabolic demand of `r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Bugula'] + log(10)*averageEstimates$s25[averageEstimates$species == 'Bugula']), 2)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^. If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies, each weighing 5 mg, would have cumulative metabolic demand of `r rounded(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Bugula'] + log(5)*averageEstimates$s25[averageEstimates$species == 'Bugula']), 2)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^ according to their new size. Thus, disturbance events change not only the size distribution of colonies but also the energy requirements of the population even if total standing biomass remains unchanged `r myCite('barneche2014ele')`.

Here we have demonstrated that the effects of temperature interact in various ways with the mass-scaling of metabolic rates in different species of colonial marine invertebrates encompassing two distinct phyla. By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age. Mass-scaling exponents vary for different species, and are not consistent across growth forms nor phylogeny. Finally, our study suggests that the often observed competitive superiority of colonial organisms over unitary ones may arise because energy availability for growth and or reproduction increases more strongly with size in colonial organisms.

@@

We thank the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C. R. White, and DP110103529 to DJM).
